\textls[-0]{\thetitle}

### 2023-03-25 canonicitymodel.tex
Help the audience parse by using background colour

In all, the whole interpretation should be considered as:
\begin{enumerate}
  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
  \item Then we apply it with two substitution,  $\GSubstExt{{\GSubstWeak{1}}}{s}$ and $\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }$
  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{...}$
\end{enumerate}

\begin{align*}
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}}(\gm_+)(\gm_+') &= ((\denotesC{\ell}(\pi_1{\gm_+})(\gm_+'[0]))[1])(\pi_2{\gm_+})(\gm_+'[1]) \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesCC{A} , \denotesCC{B}) :: \denotesC{\wsig}\\
  \denotesC{\goodWSig{\Gm}{\sub{\wsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[0]}\\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[1]}\\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm')(t) &= \CWmodel~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~t \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
  % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
  \denotesC{\goodType{\Gm}{\CaseSig{"A"}{"B"}{"R"}}{}} &= 
  \Gmodel{\GluedPi{\denotesCC{"A"}}{\GluedPi {\GluedPi{\denotesCC{"B"}}{\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}} {\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}}} \\
  % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
  \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
  RS~"nil"~R &= \denotesCC{\LSigEmp} \\
  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg 
\end{align*}
\begin{align*}
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') = \CWrec \quad &\Gsub{\denotesCC{\wsig}}{\Glued{\gm}} \\ 
  & (\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\sub{\wsig}{\gm}}{\sub{\lkg}{\gm}}{w}))\\
  &f^r\\ &\sub{t}{\gm}\\ 
  &(\denotesC{t}(\gm)(\gm')) 
\end{align*}
\begin{align*}  
  \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
  \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
  \\
  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
  & \quad 
   \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\denotesCC{\wsig}}
\end{align*}

The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
%
\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}) \\
    & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
and its eliminator $\CWrec$
\begin{align*}
  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
    & \to 
    [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    \setc{b}{\goodTerm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} \sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}})}} P~(\sub{b}{ \gm }) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
\end{align*}

>>>>

Also note that, the only non-bullet-upscripted $s$ is from the pattern matching of the interpretation of $\lsig$.

In all, the whole interpretation should be considered as:
\begin{enumerate}
  \item We first get $\denotesCC{\lkgproj{2}{\lkg}}$, the glued interpretation of the projection of the linkage 
  \item Then we apply it with two substitution,  $\Gsub{\_}{\GSubstExt{{\GSubstWeak{1}}}{s}}$ and $\Gsub{\_}{\GSubstExt{\Glued{("id")}}{ \Glued{\ell'} }}$
  \item Then we use glued interpretation of dependent tuple to combine it with $\Gpair{\Glued{\ell'}}{\_}$
\end{enumerate}

\begin{align*}
  \denotesC{\goodTerm{\Gm}{\lkgproj{1}{\lkg}}{\TyLkg{\lsig}}}(\gm)(\gm') &= \denotesC{\ell}(\gm)(\gm')[0] \quad \text{ take the first element in the tuple} \\
  \denotesC{\goodTerm{\Gm, \lsigprojT{\lsig}}{\lkgproj{2}{\lkg}}{T}}(\gm_+)(\gm_+') &= ((\denotesC{\ell}(\pi_1{\gm_+})(\gm_+'[0]))[1])(\pi_2{\gm_+})(\gm_+'[1]) \\
  \denotesC{\goodWSig{\Gm}{\WSigEmp}{0}} &= "nil" \\
  \denotesC{\goodWSig{\Gm}{\WSigAdd{\wsig}{A}{B}}{n+1}} &= (\denotesCC{A} , \denotesCC{B}) :: \denotesC{\wsig}\\
  \denotesC{\goodWSig{\Gm}{\sub{\wsig}{\gm}}{n}} & \text{ is done by point-wise/component-wise substitution} \\
  \denotesC{\goodType{\Gm}{\wsigproj{j}{1}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[0]}\\
  \denotesC{\goodType{\Gm, \wsigproj{j}{1}{\wsig}}{\wsigproj{j}{2}{\wsig}}{}} &= \Gmodel{(\text{ $j$-th element of }~\denotesC{\wsig})[1]}\\
  \denotesC{\goodWSig{\Gm}{\WSigSub{\wsig}}{n}} &= "tl"~\denotesC{\wsig}\\
  \denotesC{\goodTerm{\Gm}{\wcode{\wsig}}{\cU}}(\gm)(\gm')(t) &= \CWmodel~(\Gsub{\denotesCC{\wsig}}{\Glued{\gm}})~t \\
  \denotesC{\goodTerm{\Gm}{\Wsup{i}{\wsig}{a}{b}}{\El{\wcode{\wsig}}}}(\gm)(\gm') &= \CWsup~i~(\Gsub{\denotesCC{a}}{\Glued{\gm}})~(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}}) \\
  % CsTy~A~B~R &= {\GluedPi{A}{\GluedPi {\GluedPi{B}{\Gsub{R}{\GSubstWeak{2}}}} {\Gsub{R}{\GSubstWeak{2}}}}} \\
  \denotesC{\goodType{\Gm}{\CaseSig{"A"}{"B"}{"R"}}{}} &= 
  \Gmodel{\GluedPi{\denotesCC{"A"}}{\GluedPi {\GluedPi{\denotesCC{"B"}}{\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}} {\Gsub{\denotesCC{"R"}}{\GSubstWeak{2}}}}} \\
  % CsTy~\denotesC{A}~\denotesC{B}~\denotesC{R}\\
\end{align*}
We note that, the interpretation of $\Wsup{i}{\wsig}{a}{b}$ is defined by $\CWsup$ (defined below) will take three elements, $i$, a glued $(\Gsub{\denotesCC{a}}{\Glued{\gm}})$ and a glued $(\Gsub{\denotesCC{b}}{{\Glued{\gm^\uparrow}}})$.  As we mentioned earlier, $\Glued{\gm}$ is the equivalent form of $(\gm, \gm')$.

The interpretation for $\CaseSig{"A"}{"B"}{"R"}$ is also bullet-drowned. However, it should be read as the glued interpretation of $ \TyPi{A}{\TyPi {\TyPi{B}{\sub{R}{\SubstWeak{2}}}} {\sub{R}{\SubstWeak{2}}}} $, which is exactly from the debruinj indxed-version of the typing rule \ruleref{tyeq/casety}.

In the following interpretation, parsing might be hard, so we might use background colour to make things easier to read, and there is no meaning for the choice of colour.

\begin{align*}
  \text{we define } \RecSig{}{} &\text{ by induction on the signature, via } RS \\
  RS~"nil"~R &= \denotesCC{\LSigEmp} \\
  RS~((A, B) :: tl)~R &= \GLSigAdd{RS~tl~R}{\Glued{\pi_2}}{\GCaseSig{A}{B}{R}} \\ 
  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\
  \denotesC{\Recproj{j}{\lkg}} &= \text{ take the $j$-th field from } \lkg 
\end{align*}
\begin{align*}
  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') = \CWrec \quad &\Gsub{\denotesCC{\wsig}}{\Glued{\gm}} \\ 
  & (\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\sub{\wsig}{\gm}}{\sub{\lkg}{\gm}}{w}))\\
  &f^r\\ &\sub{t}{\gm}\\ 
  &(\denotesC{t}(\gm)(\gm')) 
\end{align*}
\begin{align*}  
  \text{ where } \quad f^r~j~\Glued{a}~b~b^c &= 
  \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
  \\
  &\texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^c)~\texttt{ in } \\
  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\GRecproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
  & \quad 
   \text{ given } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \\
   & \quad \text{ where }\Glued{R} = \denotesCC{"R"}, \Glued{B} = \Gwsigproj{j}{2}{\denotesCC{\wsig}}
\end{align*}

The semantic interpretations above are defined together with the following inductively defined indexed set $\CWmodel$ (with only one constructor $\CWsup$)
%
\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{
      \colorbox{blue!10}{(\Glued{\cdot} \Glued{,} \Gsub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}})} 
      }{
        \colorbox{green!10}{\Gsub{\GEl{\Gwcode{\Glued{\wsig}}}}{\GSubstWeak{1}}}
        }) \\
    & \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
and its eliminator $\CWrec$
\begin{align*}
  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
    & \to 
    [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\Gwsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    \setc{b}{\goodTerm{({\cdot} , 
    \colorbox{blue!10}{\sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{"id"}{{a}}}}
    ) }{b}{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}} \\ 
    & \quad \quad \to  \prod_{\Glued{\gm} \in \GSub{\Glued{\cdot}}{(\Glued{\cdot} \Glued{,} 
    \colorbox{blue!10}{\sub{\Gwsigproj{j}{2}{\Glued{\wsig}}}{\GSubstExt{\GSubstId}{\Glued{a}}}} )}} P~(\sub{b}{ \gm }) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \quad \text{ (square bracket here just to make things readable)} \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~\Glued{\wsig}~P~f~(\sub{b}{\gm})~(\Gmodel{\Glued{b}}~\Glued{\gm}))
\end{align*}
